---
title: "Reproducible document"
author: "Barath Balarasa"
format: html
---

## Basics of R

```{r}
1 + 1
```


## Setup code chunk

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <-
  read_csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |>
  rename(auc_glucose = auc_pg,
         auc_insulin = auc_ins) |>
  filter(glykemi == 0) |>
  select(-glykemi)
```

## Showing the data

```{r}
post_meal_data
```


```{r}
glimpse(post_meal_data)
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-BMI-histogram
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
@fig-BMI-histogram

## Exercise: discrete plots

```{r}
#| fig-cap: "Dist. of Group"
#| label: fig-Group
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-Group above for a cool plot!


```{r}
#| fig-cap: "Dist. of Glykemi"
#| label: fig-gly
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar(fill = "#00BE0067", color = "darkred")
```

See @fig-gly above for a cool plot!

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```

```{r}
post_meal_data |>
  #| fig-cap: "Glykemi Group"
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r}
post_meal_data |>
  #| fig-cap: "Glykemi Group extended"
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

```{r side-by-side-figs}
#| label: fig-bmi-glykemi
#| fig-cap: "BMI and Glycemia, side-by-side"
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Number of those with glycemia"
#| - "Number of those with glycemia with cool colors"
#| layout-ncol: 3

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar(fill = "#00BE0067", color = "darkred")
```

## Plotting

```{r}
#| fig-cap: "Scatterplot showing BMI against AUC of bloody glucose"
#| label: fig-bmi-auc_pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point()
```



```{r}
#| fig-cap: "Scatterplot showing BMI against AUC of bloody glucose - smooth"
#| label: fig-bmi-auc_pg-smooth
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point(color = "purple") +
  geom_smooth(fill = "#00BE0067", color = "darkred")
```


## Pushing changes to remote


## Pulling files locally


## Selecting data
```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group, -Age)

select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("OGTT"))

select(post_meal_data, contains("mass"))
```


## Renaming specific columns
```{r}
small_post_meal_data |>
  rename(auc_glucose = auc_pg)

small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```

## Select and rename tasks

# Task 1.
```{r}
post_meal_data |>
  select(contains("OGTT"))
```



# Task 2.

```{r}
post_meal_data |>
  select(contains(".."), -contains("...")) |>
  rename(
    P.glucose.minus5.OGGT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )
```

# Task 3.
basic_info <- select(post_meal_data, id, BMI, Length, Weight, Age)
rename(basic_info, Height = Length)

post_meal_data|>
  select(post_meal_data, id, BMI, Length, Weight, Age) |>
  rename(basic_info, Height = Length)



## Filtering data by row

```{r}
small_post_meal_data |>
  filter(glykemi == 1)

small_post_meal_data |>
  filter(BMI >= 25)

small_post_meal_data |>
  filter( Group == "CTR")

```

Only 4 people have hypoglycemia
7 people have a BMI greater than 25
7 people are in the CTR group 

```{r}
small_post_meal_data |>
  filter(Group == "CTR", BMI >=25)

small_post_meal_data |>
  filter(Group == "CTR" | BMI >= 25)
small_post_meal_data |>
  
```

11 people in the group CTR have a BMI higher than 11
23 people have a BMI greater than 25 or they are in the CTR Group

## About me

- Barath Balarasa
- University of Sourthern Denmark

im **a** *student*

## Simple code

```{r}
3 * 3
```

## Testing for Git

it´s cold outside
