---
title: "Reproducible document"
author: "Barath Balarasa"
format: html
---

## Basics of R

```{r}
1 + 1
```


## Setup code chunk

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  here::here("data/post-meal-insulin.csv")
)
```

## Showing the data

```{r}
post_meal_data
```


```{r}
glimpse(post_meal_data)
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-BMI-histogram
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
@fig-BMI-histogram

## Exercise: discrete plots

```{r}
#| fig-cap: "Dist. of Group"
#| label: fig-Group
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-Group above for a cool plot!


```{r}
#| fig-cap: "Dist. of Glykemi"
#| label: fig-gly
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar(fill = "#00BE0067", color = "darkred")
```

See @fig-gly above for a cool plot!

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```

```{r}
post_meal_data |>
  #| fig-cap: "Glykemi Group"
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r}
post_meal_data |>
  #| fig-cap: "Glykemi Group extended"
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

```{r side-by-side-figs}
#| label: fig-bmi-glykemi
#| fig-cap: "BMI and Glycemia, side-by-side"
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Number of those with glycemia"
#| - "Number of those with glycemia with cool colors"
#| layout-ncol: 3

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar(fill = "#00BE0067", color = "darkred")
```

## Plotting

```{r}
#| fig-cap: "Scatterplot showing BMI against AUC of bloody glucose"
#| label: fig-bmi-auc_pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point()
```



```{r}
#| fig-cap: "Scatterplot showing BMI against AUC of bloody glucose - smooth"
#| label: fig-bmi-auc_pg-smooth
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point(color = "purple") +
  geom_smooth(fill = "#00BE0067", color = "darkred")
```


# install.packages("pak")
pak::pak("r-lib/usethis")

usethis::gh_token_help()

usethis::create_github_token()


## Pushing changes to remote


## Pulling files locally






## About me

- Barath Balarasa
- University of Sourthern Denmark

im **a** *student*

## Simple code

```{r}
3 * 3
```

## Testing for Git

shit is cold outside
